---
title: "HousingProject3"
author: "Ramesh Suragam"
date: "5/11/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#Loading libraries
suppressMessages(suppressWarnings(library(tidyverse)))
suppressMessages(suppressWarnings(library(ggplot2)))
suppressMessages(suppressWarnings(library(gridExtra)))
suppressMessages(suppressWarnings(library(MASS)))
suppressMessages(suppressWarnings(library(randomForest)))
```

### **Introduction:**

### **Loading Data:**
```{r}
#Loading the data
LondonData <- suppressMessages(suppressWarnings(read_csv("data/DataScienceProj.csv")))
head(LondonData)

#Checking for correlation
M <- cor(LondonData[,-1])
head(round(M,2))
library(corrplot)
corrplot(M, method="circle", type = "upper", order="hclust", sig.level = 0.01)
```

### **Data Cleanup:**
### Convert dummies to factors
```{r}
Dummy2Factor <- function(mat,lev1="Level1") {
      mat <- as.matrix(mat)
      factor((mat %*% (1:ncol(mat))) + 1,
          labels = c(lev1, colnames(mat)))
}

Age      <- Dummy2Factor(LondonData[,5:9],"PreWW1")
Type     <- Dummy2Factor(LondonData[,10:12],"Others")
Garage   <- Dummy2Factor(LondonData[,13:14],"HardStnd")
Bedrooms <- Dummy2Factor(LondonData[,18:21],"BedOne")

MyData <- data.frame(LondonData[,c(2:4,15:17,22,23,26)],Age,Type,Garage,Bedrooms)
summary(MyData)

MyData$Tenfree <- factor(MyData$Tenfree)
MyData$CenHeat <- factor(MyData$CenHeat)
MyData$BathTwo <- factor(MyData$BathTwo)
MyData$NewPropD <- factor(MyData$NewPropD)

levels(MyData$Tenfree) <- c("no", "yes")
levels(MyData$CenHeat) <- c("no", "yes")
levels(MyData$BathTwo) <- c("no", "yes")
levels(MyData$NewPropD) <- c("no", "yes")

head(MyData)
```

### **Remove Outliers:**
```{r}
par(mfrow= c(2,2))
boxplot(MyData$Purprice)
# From boxplot we can see that purprice greater then 600000 is an outlier so we will remove that
MyData <- MyData[MyData$Purprice<600000,]

boxplot(Purprice~CenHeat,data=MyData)
boxplot(Purprice~BathTwo,data=MyData)
boxplot(Purprice~Age,data=MyData)
boxplot(Purprice~Type,data=MyData)
boxplot(Purprice~Garage,data=MyData)
boxplot(Purprice~Bedrooms,data=MyData)
par(mfrow= c(1,1))
```
### **Exploratory Analysis:**
### **Checking for price:**
```{r}
p1 <- ggplot(MyData, aes(x=Purprice)) + geom_histogram(bins = 100, color="black", fill="white")+
  geom_vline(aes(xintercept=mean(Purprice)),
            color="blue", linetype="dashed", size=1)

p2 <- ggplot(MyData, aes(x=Purprice)) + 
 geom_histogram(bins = 100, aes(y=..density..), colour="black", fill="white")+
 geom_density(alpha=.2, fill="#FF6666")

p3 <- ggplot(MyData, aes(x=log(Purprice))) + 
 geom_histogram(bins = 100, aes(y=..density..), colour="black", fill="white")+
 geom_density(alpha=.2, fill="#FF6666")



grid.arrange(p1, p2, p3, nrow=3)
```



From the plots we can see that it is skewed towards the left so we can say that to make it a normal distribution we have to apply some transformation. After applying log transformation on purprice we see that purprice is normally distributed.


```{r}
p4 <- ggplot(MyData, aes(x=log(Purprice), color=MyData$CenHeat)) +
  geom_histogram(bins = 50, fill= "white", alpha=0.5, position="identity")+
  theme(legend.position = "top")

p5 <- ggplot(MyData, aes(x=log(Purprice), color=MyData$CenHeat)) +
  geom_freqpoly()+
  theme(legend.position = "top")

p6 <- ggplot(MyData, aes(x=log(Purprice), color=MyData$Tenfree)) +
  geom_freqpoly()+
  theme(legend.position = "top")

p7 <- ggplot(MyData, aes(x=log(Purprice), color=MyData$NewPropD)) +
  geom_freqpoly()+
  theme(legend.position = "top")

p8 <- ggplot(MyData, aes(x=log(Purprice), color=MyData$Age)) +
  geom_freqpoly()+
  theme(legend.position = "top")

p9 <- ggplot(MyData, aes(x=log(Purprice), color=MyData$Type)) +
  geom_freqpoly()+
  theme(legend.position = "top")

p10 <- ggplot(MyData, aes(x=log(Purprice), color=MyData$Garage)) +
  geom_freqpoly()+
  theme(legend.position = "top")

p11 <- ggplot(MyData, aes(x=log(Purprice), color=MyData$Bedrooms)) +
  geom_freqpoly()+
  theme(legend.position = "top")

grid.arrange(p4, p5, nrow=2)
grid.arrange(p6, p7, nrow=2)
grid.arrange(p8, p9, nrow=2)
grid.arrange(p10, p11, nrow=2)
```

```{reval= FALSE}
ld_model <- lm(log(Purprice)~Tenfree+CenHeat+BathTwo+NewPropD+FlorArea+ProfPct+Age+Type+Garage+Bedrooms, data = MyData)
step <- stepAIC(ld_model, direction="both")
step$anova
```

```{r}
library(leaps)

set.seed(123)
sample <- sample(c(TRUE, FALSE), nrow(MyData), replace = T, prob = c(0.6,0.4))
train <- MyData[sample, ]
test <- MyData[!sample, ]

ld_model <- regsubsets(log(Purprice)~Tenfree+CenHeat+BathTwo+NewPropD+FlorArea+ProfPct+Age+Type+Garage+Bedrooms, data = train,nvmax = 10)
results <- summary(ld_model)
plot(ld_model,scale="adjr2")

# extract and plot results
tibble(predictors = 1:10,
       adj_R2 = results$adjr2,
       Cp = results$cp,
       BIC = results$bic) %>%
  gather(statistic, value, -predictors) %>%
  ggplot(aes(predictors, value, color = statistic)) +
  geom_line(show.legend = F) +
  geom_point(show.legend = F) +
  facet_wrap(~ statistic, scales = "free")

which.max(results$adjr2)
which.min(results$bic)
which.min(results$cp)

ld_frwd <- regsubsets(log(Purprice)~Tenfree+CenHeat+BathTwo+NewPropD+FlorArea+ProfPct+Age+Type+Garage+Bedrooms, data = train, method = "forward",nvmax = 10)
results <- summary(ld_frwd)
plot(ld_frwd,scale="adjr2")

# extract and plot results
tibble(predictors = 1:10,
       adj_R2 = results$adjr2,
       Cp = results$cp,
       BIC = results$bic) %>%
  gather(statistic, value, -predictors) %>%
  ggplot(aes(predictors, value, color = statistic)) +
  geom_line(show.legend = F) +
  geom_point(show.legend = F) +
  facet_wrap(~ statistic, scales = "free")

which.min(results$cp)

ld_bkwd <- regsubsets(log(Purprice)~Tenfree+CenHeat+BathTwo+NewPropD+FlorArea+ProfPct+Age+Type+Garage+Bedrooms, data = train, method = "backward",nvmax = 10)
results <- summary(ld_bkwd)
plot(ld_bkwd,scale="adjr2")

# extract and plot results
tibble(predictors = 1:10,
       adj_R2 = results$adjr2,
       Cp = results$cp,
       BIC = results$bic) %>%
  gather(statistic, value, -predictors) %>%
  ggplot(aes(predictors, value, color = statistic)) +
  geom_line(show.legend = F) +
  geom_point(show.legend = F) +
  facet_wrap(~ statistic, scales = "free")

which.min(results$cp)

coef(ld_model,10)
coef(ld_frwd,10)
coef(ld_bkwd,10)

test_m <- model.matrix(log(Purprice) ~Tenfree+CenHeat+BathTwo+NewPropD+FlorArea+ProfPct+Age+Type+Garage+Bedrooms, data = test)

validation_errors <- vector("double", length = 10)

for(i in 1:10) {
  coef_x <- coef(ld_model, id = i)                     # extract coefficients for model size i
  pred_x <- test_m[ , names(coef_x)] %*% coef_x           # predict salary using matrix algebra
  validation_errors[i] <- mean((test$Purprice - pred_x)^2)  # compute test error btwn actual & predicted salary
}
plot(validation_errors, type = "b")

for(i in 1:10) {
  coef_x <- coef(ld_frwd, id = i)                     # extract coefficients for model size i
  pred_x <- test_m[ , names(coef_x)] %*% coef_x           # predict salary using matrix algebra
  validation_errors[i] <- mean((test$Purprice - pred_x)^2)  # compute test error btwn actual & predicted salary
}
plot(validation_errors, type = "b")

for(i in 1:10) {
  coef_x <- coef(ld_bkwd, id = i)                     # extract coefficients for model size i
  pred_x <- test_m[ , names(coef_x)] %*% coef_x           # predict salary using matrix algebra
  validation_errors[i] <- mean((test$Purprice - pred_x)^2)  # compute test error btwn actual & predicted salary
}

# plot validation errors
plot(validation_errors, type = "b")

# ld_model <- lm(Purprice~FlorArea+BathTwo+CenHeat+Type,data = MyData)
# # ld_model <- lm(Purprice~FlorArea+BathTwo+CenHeat+Type+Easting+Bedrooms+Age,data = MyData)
# summary(ld_model)

predict.regsubsets <- function(object, newdata, id ,...) {
  form <- as.formula(object$call[[2]]) 
  mat <- model.matrix(form, newdata)
  coefi <- coef(object, id = id)
  xvars <- names(coefi)
  mat[, xvars] %*% coefi
}

k <- 10
set.seed(1)
folds <- sample(1:k, nrow(MyData), replace = TRUE)
cv_errors <- matrix(NA, k, 15, dimnames = list(NULL, paste(1:15)))

for(j in 1:k) {
  
  # perform best subset on rows not equal to j
  ld_model <- regsubsets(log(Purprice) ~Tenfree+CenHeat+BathTwo+NewPropD+FlorArea+ProfPct+Age+Type+Garage+Bedrooms, MyData[folds != j, ], nvmax = 15)
  
  # perform cross-validation
  for( i in 1:15) {
    pred_x <- predict.regsubsets(ld_model, MyData[folds == j, ], id = i)
    cv_errors[j, i] <- mean((MyData$Purprice[folds == j] - pred_x)^2)
    }
}

mean_cv_errors <- colMeans(cv_errors)

plot(mean_cv_errors, type = "b")

final_best <- regsubsets(log(Purprice) ~Tenfree+CenHeat+BathTwo+NewPropD+FlorArea+ProfPct+Age+Type+Garage+Bedrooms, data = MyData , nvmax = 15)
coef(final_best, 15)
```

```{r}
fit <- randomForest(log(Purprice) ~Tenfree+CenHeat+BathTwo+NewPropD+FlorArea+ProfPct+Age+Type+Garage+Bedrooms,
          data = train,importance=TRUE,ntree=60)

importance.features <- tibble::rownames_to_column(data.frame(fit$importance[,c(1)]))
colnames(importance.features) <- c("rowname", "value")

ggplot(importance.features, aes(x = reorder(rowname, -value), y = value)) +
  geom_bar(stat = "identity", position = "dodge", fill="#E69F00", colour="black") +
  xlab("Feature") + ylab("Count") + ggtitle("Importance of a feature: Simple Random Forest classifier") +
  coord_flip()
```

```{r}
library(classInt)
library(RColorBrewer)

nClass = 10
Palette <- rev(brewer.pal(nClass,"Spectral"))
Classes <- classIntervals(MyData$Purprice,nClass,"quantile")
Colours <- findColours(Classes,Palette)
plot(MyData$Easting,MyData$Northing,pch=16,cex=0.25,col=Colours,asp=1)
```

### Geography - look at trends with linear and quadratic trend surfaces
```{r}
x <- MyData$Easting/1000
y <- MyData$Northing/1000
m.tr1 <- lm(Purprice~x+y,data=MyData)
AIC(m.tr1)
m.tr2 <- lm(Purprice~x+y+I(x^2)+I(y^2)+I(x*y),data=MyData)
AIC(m.tr2)
summary(m.tr1) # lower prices as we move east, slightly lower as w move south
summary(m.tr2) # lower AIC # higher price as we move west
stepAIC(m.tr2)
```

### Explore variation by borough  - first load the data
```{r}
library(rgdal)
library(rgeos)
LB <- readOGR(dsn="LondonBoroughs",layer="LondonBoroughs",stringsAsFactors=FALSE)  # Boroughs
LH <- SpatialPointsDataFrame(MyData[,1:2],MyData)                     # Houses
proj4string(LH) <- CRS(proj4string(LB))                               # copy CRS
plot(LB)
points(LH,pch=16,cex=0.5)
box()
```

### Add Brough names to data  - explore by type and borough - we'll need to do an overlay
```{r}
LHLB <- over(LH,LB)   # spatial join: points first, then polygons
dim(LHLB)
head(LHLB)            # data frame has LB attributes in LH order
MyData$Borough <- gsub(" London Boro","",LHLB$NAME)  # get the borough name

boxplot(Purprice~Borough,data=MyData)

Boroughs <- names(table(MyData$Borough))
NB <- length(Boroughs)
boxplot(log(Purprice)~Borough,data=MyData,outpch=16,outcol="red",outcex=0.75,xaxt="n")
axis(1,labels=Boroughs,at=1:NB,cex.axis=0.75,las=2)
title("Log(Price) by Borough")

boxplot(log(Purprice)~Borough,data=MyData[MyData$Type=="TypSemiD",],outpch=16,outcol="red",outcex=0.75,xaxt="n")
axis(1,labels=Boroughs,at=1:NB,cex.axis=0.75,las=2)
title("Log(Price) by Borough (Semi Detached only")

boxplot(log(Purprice)~Borough,data=MyData[MyData$Type=="TypFlat",],outpch=16,outcol="red",outcex=0.75,xaxt="n")
axis(1,labels=Boroughs,at=1:NB,cex.axis=0.75,las=2)
title("Log(Price) by Borough (Flats only")
```

### Ordered boxplot
```{r}
b.order <- rank(tapply(MyData$Purprice+runif(nrow(MyData)),MyData$Borough,median))

boxplot(Purprice~Borough,data=MyData,outpch=16,outcol="red",outcex=0.75,xaxt="n",at=b.order,ylim=c(0,500000))
axis(1,labels=Boroughs,at=b.order,cex.axis=0.75,las=2)
title("Price by Borough")


boxplot(log(Purprice)~Borough,data=MyData,outpch=16,outcol="red",outcex=0.75,xaxt="n",at=b.order)
axis(1,labels=Boroughs,at=b.order,cex.axis=0.75,las=2)
title("Log(Price) by Borough")
```

### standardsed residuals -s there a apttern
```{r}
MyData$stdres.9v <- stdres(model.9v)
boxplot(stdres.9v~Borough,data=MyData,outpch=16,outcol="red",outcex=0.75,xaxt="n")
axis(1,labels=Boroughs,at=1:NB,cex.axis=0.75,las=2)
title("Standardised Residual by Borough")

boxplot(stdres.9v~Borough,data=MyData,outpch=16,outcol="red",outcex=0.75,xaxt="n",ylim=c(-5,5))
axis(1,labels=Boroughs,at=1:NB,cex.axis=0.75,las=2)
title("Standardised Residual by Borough")
abline(h=0,lty=2)
```

### y-yhat negative : overproediction
### y-yhat positive : underprediction
```{r}
b.order.9v <- rank(tapply(MyData$stdres.9v+runif(nrow(MyData))*0.0001,MyData$Borough,median))
boxplot(stdres.9v~Borough,data=MyData,outpch=16,outcol="red",outcex=0.75,xaxt="n",at=b.order.9v,ylim=c(-5,5))
axis(1,labels=Boroughs,at=b.order.9v,cex.axis=0.75,las=2)
title("Standardised Residual by Borough")
abline(h=0,lty=2)
```

### Map of Boroughs with names
```{r}
head(LB$NAME)
Bname <- gsub(" London Boro","",LB$NAME)
xy <- coordinates(LB)
plot(LB)
text(xy[,1],xy[,2],Bname,col="blue",cex=0.5)
box()
title("London Borough Boundaries")

quickMap <- function(Var,nClass=10){
   require(classInt)
   require(RColorBrewer)
   Classes <- classIntervals(Var,nClass,method="quantile")
   Palette <- brewer.pal(nClass,"Reds")
   Colours <- findColours(Classes,Palette)
   plot(y)
   points(x.sdf2,cex=0.5,pch=16,col=Colours)
   }
```

### How about some borough specific models
```{r}
 data.frame(Bname,LB$NAME)                   # check ordering of names
 head(MyData)                                # and MyData
 NB <- length(LB)                            # number of boroughs
 results <- matrix(0,NB,2)                   # storage for borough legfel coefficients
 for(i in 1:NB) {
    m.x <- lm(Purprice~FlorArea,data=MyData[MyData$Borough == Bname[i],])
    results[i,] <- coef(m.x)
 }
rownames(results) <- Bname                   # add in names
colnames(results) <- c("Intercept","FlorArea")
print(results)
hist(results[,2])                            # look at FlorArea coefficient
boxplot(results[,2])
```

### borough levels plots with legend
```{r}
quickMap2 <- function(Var,nClass=9,dp=0,plotNames=FALSE){
   require(classInt)
   require(RColorBrewer)
   Classes <- classIntervals(Var,nClass,method="quantile",dataPrecision=dp)
   Palette <- brewer.pal(nClass,"Reds")
   Colours <- findColours(Classes,Palette)
   plot(LB,col=Colours)
   legend("bottomright",
      legend=names(attr(Colours,"table")),
      fill=attr(Colours,"palette"),
      cex=0.75,bty="n")
   box()
   if(plotNames) {
      xy <- coordinates(LB)
      text(xy[,1],xy[,2],Bname,col="blue",cex=0.5)
   }
}

quickMap2(results[,2])                    # without borough names
quickMap2(results[,2],plotNames=TRUE)     # with borough names
```

### and the residuals from the model? Plot the borough medians
```{r}
quickMap2(tapply(MyData$stdres.9v,MyData$Borough,median),plotNames=TRUE,dp=3)
```

```{r}
map1 <- ggplot(LB) +
  geom_polygon(aes(x = long, y = lat, group = group), fill = "white", colour = "black")
map1 + labs(x = "Longitude", y = "Latitude", title = "Map of Greater London with the borough boundaries")

map1 + geom_point(data = MyData, aes(x = Easting, y = Northing, colour = Age)) + scale_colour_manual(values = rainbow(14)) + labs(x = "Easting", y = "Northing", title = "Map of Greater London with the borough boundaries")

map1 + plot(LB,col = MyData$Age)
```

