---
title: "HousingProject"
author: "Ramesh Suragam, Nishant Kumar, Pawan Rakesh, Alok kumar"
date: "5/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Property Cost Project
### Read the data
```{r}
LondonData <- read.csv("data/DataScienceProj.csv",stringsAsFactors=FALSE)
print(dim(LondonData))
head(LondonData)
str(LondonData)
summary(LondonData)
```

### suspicious values for RetiPct, Unemploy

###
### large hosues costs more
###
```{r}
plot(LondonData[,c("FlorArea","Purprice")],pch=16,cex=0.5)
lines(lowess(LondonData[,c("FlorArea","Purprice")]),col="red")
```

### Convert dummies to factors
###  - more convenient for modelling
```{r}
Dummy2Factor <- function(mat,lev1="Level1") {
      mat <- as.matrix(mat)
      factor((mat %*% (1:ncol(mat))) + 1,
          labels = c(lev1, colnames(mat)))
}

Age      <- Dummy2Factor(LondonData[,5:9],"PreWW1")
Type     <- Dummy2Factor(LondonData[,10:12],"Others")
Garage   <- Dummy2Factor(LondonData[,13:14],"HardStnd")
Bedrooms <- Dummy2Factor(LondonData[,18:21],"BedOne")

MyData <- data.frame(LondonData[,c(2:4,15:17,22,23,26)],Age,Type,Garage,Bedrooms)
summary(MyData)
```

### Fit models for a single variable and look at AICs
###  - model with *lowest* AIC is closest to unknown 'true' model
```{r}
AICs <- rep(NA,10)
Models <- vector("list",10)
Vars <- colnames(MyData)[4:13]
for(i in 1:10) {
      Models[[i]] <- lm(formula(paste0("Purprice~",Vars[i])),data=MyData)
      AICs[i] <- AIC(Models[[i]])
}
print(AICs)
minAIC <- which.min(AICs)
print(AICs[minAIC])
print(Vars[minAIC])
summary(Models[[minAIC]])
```

### Evidence in favour of the model
### see Burnham and Anderson p70-78
```{r}
delta <- AICs - min(AICs)                      # differences
w     <- exp(-0.5*delta)/sum(exp(-0.5*delta))  # probabilitiies
w
```

### Next model
```{r}
AICs <- rep(NA,9)
Models <- vector("list",9)
Vars <- Vars[-minAIC]
for(i in 1:9) {
      Models[[i]] <- lm(formula(paste0("Purprice~FlorArea+",Vars[i])),data=MyData)
      AICs[i] <- AIC(Models[[i]])
}
print(AICs)
minAIC <- which.min(AICs)
print(AICs[minAIC])
print(Vars[minAIC])
summary(Models[[minAIC]])

AICs <- rep(NA,8)
Models <- vector("list",8)
Vars <- Vars[-minAIC]
for(i in 1:8) {
      Models[[i]] <- lm(formula(paste0("Purprice~FlorArea+Type+",Vars[i])),data=MyData)
      AICs[i] <- AIC(Models[[i]])
}
print(AICs)
minAIC <- which.min(AICs)
print(AICs[minAIC])
print(Vars[minAIC])
summary(Models[[minAIC]])


AICs <- rep(NA,7)
Models <- vector("list",7)
Vars <- Vars[-minAIC]
for(i in 1:7) {
      Models[[i]] <- lm(formula(paste0("Purprice~FlorArea+Type+BathTwo+",Vars[i])),data=MyData)
      AICs[i] <- AIC(Models[[i]])
}
print(AICs)
minAIC <- which.min(AICs)
print(AICs[minAIC])
print(Vars[minAIC])
summary(Models[[minAIC]])
```

### Another way
```{r}
library(MASS)

price.lm   <- lm(Purprice~FlorArea+Type+BathTwo+CenHeat+Age+Bedrooms+ProfPct, data=MyData)
price.step <- stepAIC(price.lm)
summary(price.step)


price.lm2   <- lm(Purprice~Easting+Northing+FlorArea+Type+BathTwo+CenHeat+Age+Bedrooms+ProfPct, data=MyData)
price.step2 <- stepAIC(price.lm2)
summary(price.step2)

library(classInt)
library(RColorBrewer)

nClass = 10
Palette <- rev(brewer.pal(nClass,"Spectral"))
Classes <- classIntervals(MyData$Purprice,nClass,"quantile")
Colours <- findColours(Classes,Palette)
plot(MyData$Easting,MyData$Northing,pch=16,cex=0.25,col=Colours,asp=1)
```

### Geography
```{r}
x <- MyData$Easting/1000
y <- MyData$Northing/1000
m.tr1 <- lm(Purprice~x+y,data=MyData)
AIC(m.tr1)
m.tr2 <- lm(Purprice~x+y+I(x^2)+I(y^2)+I(x*y),data=MyData)
AIC(m.tr2)
summary(m.tr1) # lower prices as we move east, slightly lower as w move south
summary(m.tr2) # lower AIC # higher price as we move west
stepAIC(m.tr2)
```

### More geography
```{r}
library(rgdal)
library(rgeos)
LB <- readOGR(dsn=".",layer="LondonBoroughs/LondonBoroughs",stringsAsFactors=FALSE)
LH <- SpatialPointsDataFrame(MyData[,1:2],MyData)
proj4string(LH) <- CRS(proj4string(LB))
plot(LB)
points(LH,pch=16,cex=0.5)
box()
```

### Add Brough names to data
```{r}
LHLB <- over(LH,LB)   # points first, then polygons
dim(LHLB)
head(LHLB)
MyData$Borough <- gsub(" London Boro","",LHLB$NAME)

boxplot(Purprice~Borough,data=MyData)
Boroughs <- names(table(MyData$Borough))
NB <- length(Boroughs)
boxplot(log(Purprice)~Borough,data=MyData,outpch=16,outcol="red",outcex=0.75,xaxt="n")
axis(1,labels=Boroughs,at=1:NB,cex.axis=0.75,las=2)
title("Log(Price) by Borough")
```

### Map of Boroughs with names
```{r}
head(LB$NAME)
Bname <- gsub(" London Boro","",LB$NAME)
xy <- coordinates(LB)
plot(LB)
text(xy[,1],xy[,2],Bname,col="blue",cex=0.5)
box()
title("London Borough Boundaries")
```