---
title: "HousingProject2"
author: "Ramesh Suragam Alok Kumar Nishant Kumar Pawan Rakesh"
date: "5/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Property Cost Project
### Read the data
```{r}
library(MASS)
LondonData <- read.csv("data/DataScienceProj.csv",stringsAsFactors=FALSE)
print(dim(LondonData))
head(LondonData)
str(LondonData)
summary(LondonData)
boxplot(LondonData$Purprice)
LondonData <- LondonData[LondonData$Purprice < 600000,]
boxplot(LondonData$Purprice)
```



### suspicious values for RetiPct, Unemploy
### large hosues costs more
```{r}
plot(LondonData[,c("FlorArea","Purprice")],pch=16,cex=0.5)
lines(lowess(LondonData[,c("FlorArea","Purprice")]),col="red")
```

### Convert dummies to factors
###  - more convenient for modelling  -
```{r}
Dummy2Factor <- function(mat,lev1="Level1") {
      mat <- as.matrix(mat)
      factor((mat %*% (1:ncol(mat))) + 1,
          labels = c(lev1, colnames(mat)))
}

Age      <- Dummy2Factor(LondonData[,5:9],"PreWW1")
Type     <- Dummy2Factor(LondonData[,10:12],"Others")
Garage   <- Dummy2Factor(LondonData[,13:14],"HardStnd")
Bedrooms <- Dummy2Factor(LondonData[,18:21],"BedOne")

MyData <- data.frame(LondonData[,c(2:4,15:17,22,23,26)],Age,Type,Garage,Bedrooms)
summary(MyData)
```

### LondonData is the original
### MyData has the factor-ed versions

### explore
```{r}
boxplot(Purprice~CenHeat,data=MyData)
boxplot(Purprice~BathTwo,data=MyData)
boxplot(Purprice~Age,data=MyData)
boxplot(Purprice~Type,data=MyData)
boxplot(Purprice~Garage,data=MyData)
boxplot(Purprice~Bedrooms,data=MyData)
```

### Fit models for a single variable and look at AICs
###  - model with *lowest* AIC is closest to unknown 'true' model
```{r}
AICs <- rep(NA,10)
Models <- vector("list",10)
Vars <- colnames(MyData)[4:13]
for(i in 1:10) {
      Models[[i]] <- lm(formula(paste0("Purprice~",Vars[i])),data=MyData)
      AICs[i] <- AIC(Models[[i]])
}
print(AICs)
minAIC <- which.min(AICs)
print(AICs[minAIC])
print(Vars[minAIC])
summary(Models[[minAIC]])
```

### have a look at the differences
```{r}
names(AICs) <- Vars                         # add names
sAICs <- sort(AICs)                         # sort into order
print(sAICs)
plot(sAICs,xaxt="n")                        # plot
axis(1,labels=names(sAICs),at=1:length(Vars),las=2,cex.axis=.75)

for(i in 2:length(Vars)){                    # compute differences
   cat(paste(names(sAICs)[i],sAICs[i]-sAICs[i-1],"\n"))
}
```

### So the model with the lowest AIC is FlorArea - most variables add a little something
```{r}
model.9v <- lm(Purprice~FlorArea+Bedrooms+Type+BathTwo+Garage+Tenfree+CenHeat+Age+ProfPct,data=MyData)
summary(model.9v)    # adj r^2 ~ .56
```

### plot simple map
```{r}
library(classInt)
library(RColorBrewer)

nClass = 10
Palette <- rev(brewer.pal(nClass,"Spectral"))
Classes <- classIntervals(MyData$Purprice,nClass,"quantile")
Colours <- findColours(Classes,Palette)
plot(MyData$Easting,MyData$Northing,pch=16,cex=0.25,col=Colours,asp=1)
```

### Geography - look at trends with linear and quadratic trend surfaces
```{r}
x <- MyData$Easting/1000
y <- MyData$Northing/1000
m.tr1 <- lm(Purprice~x+y,data=MyData)
AIC(m.tr1)
m.tr2 <- lm(Purprice~x+y+I(x^2)+I(y^2)+I(x*y),data=MyData)
AIC(m.tr2)
summary(m.tr1) # lower prices as we move east, slightly lower as w move south
summary(m.tr2) # lower AIC # higher price as we move west
stepAIC(m.tr2)
```

### Explore variation by borough  - first load the data
```{r}
library(rgdal)
library(rgeos)
LB <- readOGR(dsn="LondonBoroughs",layer="LondonBoroughs",stringsAsFactors=FALSE)  # Boroughs
LH <- SpatialPointsDataFrame(MyData[,1:2],MyData)                     # Houses
proj4string(LH) <- CRS(proj4string(LB))                               # copy CRS
plot(LB)
points(LH,pch=16,cex=0.5)
box()
```

### Add Brough names to data  - explore by type and borough - we'll need to do an overlay
```{r}
LHLB <- over(LH,LB)   # spatial join: points first, then polygons
dim(LHLB)
head(LHLB)            # data frame has LB attributes in LH order
MyData$Borough <- gsub(" London Boro","",LHLB$NAME)  # get the borough name

boxplot(Purprice~Borough,data=MyData)

Boroughs <- names(table(MyData$Borough))
NB <- length(Boroughs)
boxplot(log(Purprice)~Borough,data=MyData,outpch=16,outcol="red",outcex=0.75,xaxt="n")
axis(1,labels=Boroughs,at=1:NB,cex.axis=0.75,las=2)
title("Log(Price) by Borough")

boxplot(log(Purprice)~Borough,data=MyData[MyData$Type=="TypSemiD",],outpch=16,outcol="red",outcex=0.75,xaxt="n")
axis(1,labels=Boroughs,at=1:NB,cex.axis=0.75,las=2)
title("Log(Price) by Borough (Semi Detached only")

boxplot(log(Purprice)~Borough,data=MyData[MyData$Type=="TypFlat",],outpch=16,outcol="red",outcex=0.75,xaxt="n")
axis(1,labels=Boroughs,at=1:NB,cex.axis=0.75,las=2)
title("Log(Price) by Borough (Flats only")
```

### Ordered boxplot
```{r}
b.order <- rank(tapply(MyData$Purprice+runif(nrow(MyData)),MyData$Borough,median))

boxplot(Purprice~Borough,data=MyData,outpch=16,outcol="red",outcex=0.75,xaxt="n",at=b.order,ylim=c(0,500000))
axis(1,labels=Boroughs,at=b.order,cex.axis=0.75,las=2)
title("Price by Borough")


boxplot(log(Purprice)~Borough,data=MyData,outpch=16,outcol="red",outcex=0.75,xaxt="n",at=b.order)
axis(1,labels=Boroughs,at=b.order,cex.axis=0.75,las=2)
title("Log(Price) by Borough")
```

### standardsed residuals -s there a apttern
```{r}
MyData$stdres.9v <- stdres(model.9v)
boxplot(stdres.9v~Borough,data=MyData,outpch=16,outcol="red",outcex=0.75,xaxt="n")
axis(1,labels=Boroughs,at=1:NB,cex.axis=0.75,las=2)
title("Standardised Residual by Borough")

boxplot(stdres.9v~Borough,data=MyData,outpch=16,outcol="red",outcex=0.75,xaxt="n",ylim=c(-5,5))
axis(1,labels=Boroughs,at=1:NB,cex.axis=0.75,las=2)
title("Standardised Residual by Borough")
abline(h=0,lty=2)
```

### y-yhat negative : overproediction
### y-yhat positive : underprediction
```{r}
b.order.9v <- rank(tapply(MyData$stdres.9v+runif(nrow(MyData))*0.0001,MyData$Borough,median))
boxplot(stdres.9v~Borough,data=MyData,outpch=16,outcol="red",outcex=0.75,xaxt="n",at=b.order.9v,ylim=c(-5,5))
axis(1,labels=Boroughs,at=b.order.9v,cex.axis=0.75,las=2)
title("Standardised Residual by Borough")
abline(h=0,lty=2)
```

### Map of Boroughs with names
```{r}
head(LB$NAME)
Bname <- gsub(" London Boro","",LB$NAME)
xy <- coordinates(LB)
plot(LB)
text(xy[,1],xy[,2],Bname,col="blue",cex=0.5)
box()
title("London Borough Boundaries")

quickMap <- function(Var,nClass=10){
   require(classInt)
   require(RColorBrewer)
   Classes <- classIntervals(Var,nClass,method="quantile")
   Palette <- brewer.pal(nClass,"Reds")
   Colours <- findColours(Classes,Palette)
   plot(y)
   points(x.sdf2,cex=0.5,pch=16,col=Colours)
   }
```

### How about some borough specific models
```{r}
 data.frame(Bname,LB$NAME)                   # check ordering of names
 head(MyData)                                # and MyData
 NB <- length(LB)                            # number of boroughs
 results <- matrix(0,NB,2)                   # storage for borough legfel coefficients
 for(i in 1:NB) {
    m.x <- lm(Purprice~FlorArea,data=MyData[MyData$Borough == Bname[i],])
    results[i,] <- coef(m.x)
 }
rownames(results) <- Bname                   # add in names
colnames(results) <- c("Intercept","FlorArea")
print(results)
hist(results[,2])                            # look at FlorArea coefficient
boxplot(results[,2])
```

### borough levels plots with legend
```{r}
quickMap2 <- function(Var,nClass=9,dp=0,plotNames=FALSE){
   require(classInt)
   require(RColorBrewer)
   Classes <- classIntervals(Var,nClass,method="quantile",dataPrecision=dp)
   Palette <- brewer.pal(nClass,"Reds")
   Colours <- findColours(Classes,Palette)
   plot(LB,col=Colours)
   legend("bottomright",
      legend=names(attr(Colours,"table")),
      fill=attr(Colours,"palette"),
      cex=0.75,bty="n")
   box()
   if(plotNames) {
      xy <- coordinates(LB)
      text(xy[,1],xy[,2],Bname,col="blue",cex=0.5)
   }
}

quickMap2(results[,2])                    # without borough names
quickMap2(results[,2],plotNames=TRUE)     # with borough names
```

### and the residuals from the model? Plot the borough medians
```{r}
quickMap2(tapply(MyData$stdres.9v,MyData$Borough,median),plotNames=TRUE,dp=3)
```